---
title: "EVAnalysis"
output: html_document
date: '2022-04-09'
---

```{r setup}
library(tidyverse)
library(lubridate)
library(readxl)
library(plotly)
library(RColorBrewer)
knitr::opts_chunk$set(echo = TRUE)
```

Read the data, run a summary on the data and create two new columns Range_Miles and PriceDollars

```{r read ev_cars}
ev_cars <- read_csv("ElectricCarData_Clean.csv")
summary(ev_cars)

ev_cars <- ev_cars%>%
  mutate(Range_Miles = round(Range_Km * 0.621371,0)) %>%
  mutate(PriceDollars = round(PriceEuro * 1.09))

```
Get the Top 10 EVs by range in miles and create a Bar graph showing the Brand, Model and Range in Miles

```{r top 10 cars by range}
ev_cars_top10 <- ev_cars[order(-ev_cars$Range_Miles),]

ev_cars_top10 <- ev_cars_top10[1:10,]

ev_cars_top10 %>%
  ggplot(aes(fill =Model, y = Range_Miles, x = Brand)) + 
  geom_bar(position="dodge", stat="identity") +
  xlab("Electric Vehicles") +
  ylab("Range in Miles") +
  ggtitle("Range in Miles for Top 10 EV Models") +
  theme(plot.title = element_text(hjust = 0.5),axis.text.x = element_text(angle = 45)) +
  scale_fill_brewer(type = "qual", palette = 3)

  
```

Calculate the mean efficiency of each Brand and generate a Bar Plot showing Top 20 Brands by Mean Efficiency
```{r Brand by Mean Efficiency}
ev_cars_mean_effi_brand <- ev_cars %>%
  group_by(Brand) %>%
  summarise(mean_efficiency = round(mean(Efficiency_WhKm),0))

ev_cars_mean_effi_brand <- ev_cars_mean_effi_brand[order(-ev_cars_mean_effi_brand$mean_efficiency),]
ev_cars_mean_effi_brand %>%
  mutate(Brand = fct_rev(fct_reorder(Brand, mean_efficiency))) %>%
  head(n = 20) %>%
  ggplot(aes(Brand, mean_efficiency,fill = Brand)) +
  geom_bar(position="dodge", stat="identity") +
  xlab("Brand") +
  ylab("Mean Efficiency") +
  ggtitle("Top 20 Brands by Mean Efficiency") +
  theme(plot.title = element_text(hjust = 0.5),axis.text.x = element_text(angle = 45), legend.position="none") 
  

```



Generate a Boxplots showing the Price by Body Style
```{r Boxplot}
ev_cars %>%
  ggplot(aes(BodyStyle, PriceDollars, fill = BodyStyle)) +
  geom_boxplot() +
  xlab("Body Style") +
  ylab("Price in Dollars") +
  ggtitle("Price by Body Style") +
  theme(plot.title = element_text(hjust = 0.5), legend.position="none") 
 
ggplotly()

```

As Tesla is one of the most well known EV brands in the market, below is a bubble graph showing all Tesla models by Price, Range and Body Style.
```{r Tesla}
ev_cars %>%
  filter(Brand == "Tesla") %>%
  ggplot(aes(x = Model, y = PriceDollars, size = Range_Miles, color = BodyStyle)) +
  geom_point(alpha = 0.5) +
  scale_size(range = c(1, 15)) +
  xlab("Tesla Models") +
  ylab("Price in Dollars") +
  ggtitle("A Closer Look at the Different Tesla Models") + 
  theme(plot.title = element_text(hjust = 0.5),axis.text.x = element_text(angle = 90))

ggplotly()
  
```


